---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# __genNetDem__

**genNetDem** is an R package that that generates longitudinal social network and/or capture-recapture datasets with known underlying properties. 

## Bugs and Suggestions

It is very difficult to develop a package that is robust to all types of datasets and user requirements. But we want to help. Please report any bugs or suggestions in the [Issues](https://github.com/matthewsilk/NETDEM/issues).

## Installation

**genNetDem** can be easily downloaded and installed from GitHub using `remotes::install_github()`. The vignettes available with the package can be downloaded by adding `build_vignettes = TRUE`.

```{r install, message=FALSE, eval = FALSE}
# install CMRnet
remotes::install_github("matthewsilk/NETDM/genNetDem", build_vignettes = TRUE)
```

## Getting started

- For an introduction to using **genNetDem**, see `vignette("using genNetDem")`
